---
title: "Standardized Mean Differences (SMD)"
author: "Shawn Garbett"
date: '`r Sys.Date()`'
output:
  pdf_document: null
  html_document: null
header-includes:
- \usepackage[LGR,T1]{fontenc}
- \usepackage[utf8]{inputenc}
- \usepackage{textgreek}
- \usepackage{float}
- \usepackage[x11names,dvipsnames,table]{xcolor}
- \usepackage{boldline}
- \usepackage{colortbl}
- \usepackage{hhline}
- \usepackage{longtable}
- \pdfminorversion=5
- \pdfcompresslevel=9
- \pdfobjcompresslevel=2
vignette: |
  %\VignetteIndexEntry{Tangram Standardized Mean Differences (SMD)}
  %\VignetteEngine{knitr::rmarkdown}
---

<style type="text/css">

body, td {
   font-size: 12px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 10px
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Purpose

To create a SMD statistic in the style of tableone inside of tangram. See '[Using standardized mean differences](https://CRAN.R-project.org/package=tableone/vignettes/smd.html)' by Kazuki Yoshida for original work and description of method. This is a reproduction of the original document using `tangram`.

## Load Packages and Data

The right heart catheterization dataset is available at [http://biostat.mc.vanderbilt.edu/wiki/Main/DataSets](http://biostat.mc.vanderbilt.edu/wiki/Main/DataSets) . This dataset was originally used in Connors et al. JAMA 1996;276:889-897, and has been made publicly available.

```{r, results=FALSE}
library(Matching)
library(tangram)
## Right heart cath dataset
rhc <- read.csv("http://biostat.mc.vanderbilt.edu/wiki/pub/Main/DataSets/rhc.csv")
```

## Unmatched table
```{r, results='asis'}
vars <- c("age","sex","race","edu","income","ninsclas","cat1","das2d3pc","dnr1",
          "ca","surv2md1","aps1","scoma1","wtkilo1","temp1","meanbp1","resp1",
          "hrt1","pafi1","paco21","ph1","wblc1","hema1","sod1","pot1","crea1",
          "bili1","alb1","resp","card","neuro","gastr","renal","meta","hema",
          "seps","trauma","ortho","cardiohx","chfhx","dementhx","psychhx",
          "chrpulhx","renalhx","liverhx","gibledhx","malighx","immunhx",
          "transhx","amihx")

formula <- paste0("swang1~", paste0(vars, collapse="+"))

tangram(formula, rhc, "unmatched", smd, "Unmatched", style="nejm", pct_width=0.5, test=TRUE)
```

## Propensity Score Matching
```{r, results='asis'}
## Fit model
psModel <- glm(formula = formula,
               family  = binomial(link = "logit"),
               data    = rhc)

## Predicted probability of being assigned to RHC
rhc$pRhc <- predict(psModel, type = "response")
## Predicted probability of being assigned to no RHC
rhc$pNoRhc <- 1 - rhc$pRhc

## Predicted probability of being assigned to the
## treatment actually assigned (either RHC or no RHC)
rhc$pAssign <- NA
rhc$pAssign[rhc$swang1 == "RHC"]    <- rhc$pRhc[rhc$swang1   == "RHC"]
rhc$pAssign[rhc$swang1 == "No RHC"] <- rhc$pNoRhc[rhc$swang1 == "No RHC"]
## Smaller of pRhc vs pNoRhc for matching weight
rhc$pMin <- pmin(rhc$pRhc, rhc$pNoRhc)

listMatch <- Match(Tr       = (rhc$swang1 == "RHC"),      # Need to be in 0,1
                   ## logit of PS,i.e., log(PS/(1-PS)) as matching scale
                   X        = log(rhc$pRhc / rhc$pNoRhc),
                   ## 1:1 matching
                   M        = 1,
                   ## caliper = 0.2 * SD(logit(PS))
                   caliper  = 0.2,
                   replace  = FALSE,
                   ties     = TRUE,
                   version  = "fast")
## Extract matched data
rhcMatched <- rhc[unlist(listMatch[c("index.treated","index.control")]), ]

tangram(formula, rhcMatched, "matched", smd, "Propensity Score Matched", style="nejm", pct_width=0.5)
```

## Propensity Score Matching Weight
```{r, results='asis'}
tangram(formula, rhc, "weighted", smd,
        "Propensity Score Weighted",
        style="nejm",
        weight=rhc$pMin/rhc$pAssign, pct_width=0.5)
```
